---
title: "[__cellcontacts__](https://github.com/juliaquach02/cellcontacts)"
output: md_document
author: "Julia Quach"
date: "2023-04-29"
bibliography: rmarkdown.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

--------------------------------------------------------------------------------

# __About__

--------------------------------------------------------------------------------

This R package was written to process imaging files acquired from live cell imaging of T cells and tumor cells. It requires the prior analysis of the live cell imaging films with a cell tracking tool, e.g., [TrackMate](https://imagej.net/plugins/trackmate/) or [Imaris](https://imaris.oxinst.com/products/imaris-for-tracking). It offers functions to compute cell-cell distances, cell-cell contacts, and characteristics of cell-cell interaction. Moreover, it allows to connect the results to immunological staining results.
 
--------------------------------------------------------------------------------

# __Goal of the project__

The goal of this R package is to enable the quantitative analysis of T cell-tumor cell interaction in live cell imaging films.

--------------------------------------------------------------------------------

# __Installation__ 

The latest development version of cellcontacts can be installed from GitHub:

```{r, eval = FALSE}
devtools::install_github( "juliaquach02/cellcontacts" )
```
The package can be loaded via:

```{r, eval = FALSE}
library( cellcontacts )
```

--------------------------------------------------------------------------------

# __Example__

The package was built to analyze 2D live-cell imaging movies. Before using the package,
the cells need to segmented and tracked, for example using the Fiji plugin 
[TrackMate](https://imagej.net/plugins/trackmate/) by Tinevez at al [@RN293]. 

The cellcontacts package processes 

* ROIs representing segmented cells, which are saved as .roi-files in a zipped folder, and 
* cell tracks, which are saved as .csv file with the columns "Label", "X", "Y", "Frame".

## __Prior analysis__

TrackMate is a modular tool that enables the segmentation and tracking of cells 
via common segmentation tools, e.g. [StarDist](https://imagej.net/plugins/stardist), 
and tracking tools, e.g. the [LAP tracker](). 

This package requires separate segmentation and tracking of tumor and T cells.
After segmentation and tracking, ROIs should be renamed in the TrackMate GUI 
by their spot name with incremented index. Afterwards, they can be exported 
to the ROI manager, from where they can be saved as zipped folder.In addition, cell tracks should be saved from the TrackMate GUI as .csv file.

ADD SCREENSHOTS

## __Setting directories and loading the data__

Before starting the analysis, source directories need to be set.


Also, destination directories need to be set.

To load tracks from a .csv-file, we simply use the `read.table()` function. To load the cell ROIs, we use the `read.ijzip()` function from the [RImageJROI](https://cran.r-project.org/web/packages/RImageJROI/index.html) package.
As the number of ROIs can be very large, we might need to split the zipped ROIs into multiple zipped subfolders and load the folders in parallel using the [future](https://cran.r-project.org/web/packages/future/index.html) package. To save the ROIs into separate subfolders, an ImageJ macro is available [here]().

## __Analysis__


We follow multiple steps to compute cell-cell distances and to filter for cell-cell contacts.

### 1. Matching track names to their frame number

### 2. Putting data into hash maps

### 3. Computing cell-cell distances

### 4. Filtering for cell-cell contacts

### 5. Export the results for visualization and validation

--------------------------------------------------------------------------------
# __References__
